---
title: "APHS_nonbinary"
author: "Varun Warrier"
date: "30 May 2019"
output:
  html_document: default
  pdf_document: default
---

```{r}
library(ggplot2)
library(epitools)
library(ggplot2)
library(data.table)
library(logistf)
load("APHS.RData")
```

## Autism: case-control


###Calculates odds ratio by median-unbiased estimation (mid-p), conditional maximum likelihood estimation (Fisher), unconditional maximum likelihood estimation (Wald), and small sample adjustment (small). Confidence intervals are calculated using exact methods (mid-p and Fisher), normal approximation (Wald), and normal approximation with small sample adjustment (small). This function expects the following table struture:

                    disease=0   disease=1
    exposed=0 (ref)    n00         n01
    exposed=1          n10         n11	
    exposed=2          n20         n21
    exposed=3          n30         n31
  
The reason for this is because each level of exposure is compared to the reference level. If you are providing a 2x2 table the following table is preferred:
                    disease=0   disease=1
    exposed=0 (ref)    n00         n01
    exposed=1          n10         n11	
  ###

                    Controls      Cases
    exposed=Males    n00         n01
    exposed=Nonbin   n10         n11	




```{r case-control chi-square}
males = subset(APHS, Binary_nonbinary == "Binary_Male")
autism_males = subset(males, Autism.Y.N == "1")
control_males = subset(males, Autism.Y.N == "0")

females = subset(APHS, Binary_nonbinary == "Binary_Female")
autism_females = subset(females, Autism.Y.N == "1")
control_females = subset(females, Autism.Y.N == "0")

nonbinary = subset(APHS, Binary_nonbinary == "Nonbinary")
autism_nonbinary = subset(nonbinary, Autism.Y.N == "1")
control_nonbinary = subset(nonbinary, Autism.Y.N == "0")



sex_3_way = matrix(c(nrow(control_males), nrow(autism_males), nrow(control_females), nrow(autism_females), nrow(control_nonbinary), nrow(autism_nonbinary) ), ncol = 2, byrow = TRUE)

colnames(sex_3_way) = c("Controls", "Cases")

rownames(sex_3_way) = c("Males", "Females", "Nonbinary")

chisq.test(sex_3_way)


#############


male_nonbinary = matrix(c(nrow(control_males), nrow(autism_males), nrow(control_nonbinary), nrow(autism_nonbinary) ), ncol = 2, byrow = TRUE)

colnames(male_nonbinary) = c("Controls", "Cases")

rownames(male_nonbinary) = c("Males", "Nonbinary")

male_nonbinary

chisq.test(male_nonbinary)
oddsratio(male_nonbinary)


##########

female_nonbinary = matrix(c(nrow(control_females), nrow(autism_females), nrow(control_nonbinary), nrow(autism_nonbinary) ), ncol = 2, byrow = TRUE)

colnames(female_nonbinary) = c("Controls", "Cases")

rownames(female_nonbinary) = c("Females", "Nonbinary")

female_nonbinary

chisq.test(female_nonbinary)
oddsratio(female_nonbinary)



##########

binary_nonbinary = matrix(c(nrow(control_females) + nrow(control_males), nrow(autism_females) + nrow(autism_males), nrow(control_nonbinary), nrow(autism_nonbinary) ), ncol = 2, byrow = TRUE)

colnames(binary_nonbinary) = c("Controls", "Cases")

rownames(binary_nonbinary) = c("Binary", "Nonbinary")

binary_nonbinary

chisq.test(binary_nonbinary)
oddsratio(binary_nonbinary)


### Regressing age effects


summary(glm(Autism.Y.N ~  relevel(as.factor(Binary_nonbinary), ref = "Nonbinary") + scale(Age) + Education, data = APHS, family = "binomial"))

summary(glm(Autism.Y.N ~  binary2 + scale(Age) + Education, data = APHS, family = "binomial"))



```



## Create plots

```{R, Plot}

ggplot(APHS, aes(x=Age, colour=Binary_nonbinary)) +   geom_density() + theme_classic() + xlab("Age")

summary(aov(Age ~ Binary_nonbinary, data = APHS))


APHS$edu2 = ifelse(APHS$Education == "University Postgraduate level qualifications (MA, MSc, PhD, Certificate, etc.)", "1 - Postgraduate", 0)
APHS$edu2 = ifelse(APHS$Education == "University Undergraduate level qualifications (BA, BSc, etc.)", "2 - Undergraduate", APHS$edu2)
APHS$edu2 = ifelse(APHS$Education == "Further vocational qualifications", "3 - Vocational", APHS$edu2)
APHS$edu2 = ifelse(APHS$Education == "Secondary School/ High School level qualifications", "4 - Highschool", APHS$edu2)
APHS$edu2 = ifelse(APHS$Education == "No formal qualifications", "5 - No Highschool", APHS$edu2)
APHS$edu2 = ifelse(APHS$Education == "", "Missing", APHS$edu2)

summary_edu = setDT(APHS)[, list(count=.N) , list(Binary_nonbinary, edu2)]

summary_edu


ggplot(APHS, aes(Binary_nonbinary, fill = edu2)) + geom_bar(position = 'fill') + theme_classic()


```

