---
title: "Non_binary_CARD"
author: "Varun Warrier"
date: "21 April 2019"
output:
  html_document: default
  pdf_document: default
---

# Libraries and read in the data
```{r setup, include=FALSE}
library(ggplot2)
library(epitools)
library(ggplot2)
library(data.table)
load("IMAGEnonbinary.RData")
```

## Autism: case-control


###Calculates odds ratio by median-unbiased estimation (mid-p), conditional maximum likelihood estimation (Fisher), unconditional maximum likelihood estimation (Wald), and small sample adjustment (small). Confidence intervals are calculated using exact methods (mid-p and Fisher), normal approximation (Wald), and normal approximation with small sample adjustment (small). This function expects the following table struture:

                    disease=0   disease=1
    exposed=0 (ref)    n00         n01
    exposed=1          n10         n11	
    exposed=2          n20         n21
    exposed=3          n30         n31
  
The reason for this is because each level of exposure is compared to the reference level. If you are providing a 2x2 table the following table is preferred:
                    disease=0   disease=1
    exposed=0 (ref)    n00         n01
    exposed=1          n10         n11	
  ###

                    Controls      Cases
    exposed=Males    n00         n01
    exposed=Nonbin   n10         n11	




```{r case-control chi-square}

CARD2$key = ifelse(CARD2$Gender_final == "Male" & CARD2$ASCDiagnosis == "Y" , "autism_males", "other")
CARD2$key = ifelse(CARD2$Gender_final == "Male" & CARD2$ASCDiagnosis == "N" , "control_males", CARD2$key)
CARD2$key = ifelse(CARD2$Gender_final == "Female" & CARD2$ASCDiagnosis == "N" , "control_females", CARD2$key)
CARD2$key = ifelse(CARD2$Gender_final == "Female" & CARD2$ASCDiagnosis == "Y" , "autism_females", CARD2$key)
CARD2$key = ifelse(CARD2$Gender_final == "Nonbinary" & CARD2$ASCDiagnosis == "Y" , "autism_nonbinary", CARD2$key)
CARD2$key = ifelse(CARD2$Gender_final == "Nonbinary" & CARD2$ASCDiagnosis == "N" , "control_nonbinary", CARD2$key)

autism_males = subset(CARD2, key == "autism_males")
autism_females = subset(CARD2, key == "autism_females")
autism_nonbinary = subset(CARD2, key == "autism_nonbinary")
control_males = subset(CARD2, key == "control_males")
control_females = subset(CARD2, key == "control_females")
control_nonbinary = subset(CARD2, key == "control_nonbinary")

sex_3_way = matrix(c(nrow(control_males), nrow(autism_males), nrow(control_females), nrow(autism_females), nrow(control_nonbinary), nrow(autism_nonbinary) ), ncol = 2, byrow = TRUE)

colnames(sex_3_way) = c("Controls", "Cases")

rownames(sex_3_way) = c("Males", "Females", "Nonbinary")

chisq.test(sex_3_way)

oddsratio(sex_3_way)



#############


male_nonbinary = matrix(c(nrow(control_males), nrow(autism_males), nrow(control_nonbinary), nrow(autism_nonbinary) ), ncol = 2, byrow = TRUE)

colnames(male_nonbinary) = c("Controls", "Cases")

rownames(male_nonbinary) = c("Males", "Nonbinary")

male_nonbinary

chisq.test(male_nonbinary)
oddsratio(male_nonbinary)


##########

female_nonbinary = matrix(c(nrow(control_females), nrow(autism_females), nrow(control_nonbinary), nrow(autism_nonbinary) ), ncol = 2, byrow = TRUE)

colnames(female_nonbinary) = c("Controls", "Cases")

rownames(female_nonbinary) = c("Females", "Nonbinary")

female_nonbinary

chisq.test(female_nonbinary)
oddsratio(female_nonbinary)



##########

binary_nonbinary = matrix(c(nrow(control_females) + nrow(control_males), nrow(autism_females) + nrow(autism_males), nrow(control_nonbinary), nrow(autism_nonbinary) ), ncol = 2, byrow = TRUE)

colnames(binary_nonbinary) = c("Controls", "Cases")

rownames(binary_nonbinary) = c("Binary", "Nonbinary")

binary_nonbinary

chisq.test(binary_nonbinary)
oddsratio(binary_nonbinary)


### Regressing age effects

CARD2$autism2 = ifelse(CARD2$ASCDiagnosis == "Y", 1, 0)

summary(glm(autism2 ~ Gender_3 + scale(AQ_Age) + Education_recode + Category, data = CARD2, family = "binomial"))
summary(glm(autism2 ~ relevel(as.factor(Gender_final), ref = "Nonbinary") + scale(AQ_Age) + Education_recode + Category, data = CARD2, family = "binomial" ))



```



## Let's do suspected autism now

```{R, suspected autism}

####Suspected autism


suspectall = matrix(c(589, 216, 490, 97, 10, 16 ), ncol = 2, byrow = TRUE)

colnames(suspectall) = c("Controls", "Autism_suspect")

rownames(suspectall) = c("Males", "Females", "Nonbinary")

suspectall

chisq.test(suspectall)

oddsratio(suspectall)


### Suspect_males

suspectall_males = matrix(c(589, 216,  10, 16 ), ncol = 2, byrow = TRUE)

colnames(suspectall_males) = c("Controls", "Autism_suspect")

rownames(suspectall_males) = c("Males",  "Nonbinary")

suspectall_males

chisq.test(suspectall_males)

oddsratio(suspectall_males)



### Suspect_females

suspectall_females = matrix(c(490, 97,  10, 16 ), ncol = 2, byrow = TRUE)

colnames(suspectall_females) = c("Controls", "Autism_suspect")

rownames(suspectall_females) = c("Females",  "Nonbinary")

suspectall_females

chisq.test(suspectall_females)

oddsratio(suspectall_females)



### Suspect_binary

suspectall_binary = matrix(c(490 + 589 , 97 + 216,  10, 16 ), ncol = 2, byrow = TRUE)

colnames(suspectall_binary) = c("Controls", "Autism_suspect")

rownames(suspectall_binary) = c("Binary",  "Nonbinary")

suspectall_binary

chisq.test(suspectall_binary)

oddsratio(suspectall_binary)



```

## Autistic traits regression

```{R, Autistic traits}

summary(lm(scale(AQ_TestScore) ~ as.factor(ASCDiagnosis) + Education_recode + Gender_final + scale(AQ_Age) + Category, data = CARD2))

summary(lm(scale(AQ_TestScore) ~ as.factor(ASCDiagnosis) + Education_recode + Gender_3 + scale(AQ_Age) + Category, data = CARD2))


```

## Make plot

```{R, Plot}

ggplot(CARD2, aes(x=AQ_TestScore, colour=key)) +   geom_density() + theme_classic() + xlab("AQ_full")


summary_edu = setDT(CARD2)[, list(count=.N) , list(Gender_final, Education_recode)]

summary_edu

ggplot(CARD2, aes(Gender_final, fill = Education_recode)) + geom_bar(position = 'fill') + theme_classic()

ggplot(CARD2, aes(x=AQ_Age, colour=Gender_final)) +   geom_density() + theme_classic() + xlab("AQ_full")



```

