---
title: "MU_dataset analysis"
author: "Varun Warrier"
date: "22 April 2019"
output: html_document
---

Load data and libraries
```{r setup, include=FALSE}
library(ggplot2)
library(epitools)
library(ggplot2)
library(data.table)
load("MUnonbinary.RData")



MU2$education = ifelse(MU2$education_1 == "1", "5 - No high school", 0)
MU2$education = ifelse(MU2$education_1 == "2", "4 - High school", MU2$education)
MU2$education = ifelse(MU2$education_1 == "3", "2 - Undergraduate", MU2$education)
MU2$education = ifelse(MU2$education_1 == "4", "1 - Postgraduate", MU2$education)
MU2$education = ifelse(MU2$education_1 == "5", "3 - Still studying", MU2$education)
MU2$education = ifelse(MU2$education_1 == "NA", "Missing", MU2$education)
MU2$autism = ifelse(MU2$autism_1 == "2", 0, 1)
MU2$binary = ifelse(MU2$sex_1 < 3, "binary", "nonbinary")
MU2$sex2 = ifelse(MU2$sex_1 == "1", "males", 0)
MU2$sex2 = ifelse(MU2$sex_1 == "2", "females", MU2$sex2)
MU2$sex2 = ifelse(MU2$sex_1 > 2, "nonbinary", MU2$sex2)
```

## R Markdown
## Autism: case-control


###Calculates odds ratio by median-unbiased estimation (mid-p), conditional maximum likelihood estimation (Fisher), unconditional maximum likelihood estimation (Wald), and small sample adjustment (small). Confidence intervals are calculated using exact methods (mid-p and Fisher), normal approximation (Wald), and normal approximation with small sample adjustment (small). This function expects the following table struture:

                    disease=0   disease=1
    exposed=0 (ref)    n00         n01
    exposed=1          n10         n11	
    exposed=2          n20         n21
    exposed=3          n30         n31
  
The reason for this is because each level of exposure is compared to the reference level. If you are providing a 2x2 table the following table is preferred:
                    disease=0   disease=1
    exposed=0 (ref)    n00         n01
    exposed=1          n10         n11	
  ###

                    Controls      Cases
    exposed=Males    n00         n01
    exposed=Nonbin   n10         n11	


```{r cars}
autism_males = subset(MU2, autism_1 == "1" & sex_1 == "1")
control_males = subset(MU2, autism_1 == "2" & sex_1 == "1")
autism_females = subset(MU2, autism_1 == "1" & sex_1 == "2")
control_females = subset(MU2, autism_1 == "2" & sex_1 == "2")
autism_nonbinary = subset(MU2, autism_1 == "1" & binary == "nonbinary")
control_nonbinary = subset(MU2, autism_1 == "2" & binary == "nonbinary")

sex_3_way = matrix(c(nrow(control_males), nrow(autism_males), nrow(control_females), nrow(autism_females), nrow(control_nonbinary), nrow(autism_nonbinary) ), ncol = 2, byrow = TRUE)

colnames(sex_3_way) = c("Controls", "Cases")

rownames(sex_3_way) = c("Males", "Females", "Nonbinary")

chisq.test(sex_3_way)

oddsratio(sex_3_way)



#############


male_nonbinary = matrix(c(nrow(control_males), nrow(autism_males), nrow(control_nonbinary), nrow(autism_nonbinary) ), ncol = 2, byrow = TRUE)

colnames(male_nonbinary) = c("Controls", "Cases")

rownames(male_nonbinary) = c("Males", "Nonbinary")

male_nonbinary

chisq.test(male_nonbinary)
oddsratio(male_nonbinary)


##########

female_nonbinary = matrix(c(nrow(control_females), nrow(autism_females), nrow(control_nonbinary), nrow(autism_nonbinary) ), ncol = 2, byrow = TRUE)

colnames(female_nonbinary) = c("Controls", "Cases")

rownames(female_nonbinary) = c("Females", "Nonbinary")

female_nonbinary

chisq.test(female_nonbinary)
oddsratio(female_nonbinary)



##########

binary_nonbinary = matrix(c(nrow(control_females) + nrow(control_males), nrow(autism_females) + nrow(autism_males), nrow(control_nonbinary), nrow(autism_nonbinary) ), ncol = 2, byrow = TRUE)

colnames(binary_nonbinary) = c("Controls", "Cases")

rownames(binary_nonbinary) = c("Binary", "Nonbinary")

binary_nonbinary

chisq.test(binary_nonbinary)
oddsratio(binary_nonbinary)




```

## Regression

Let's do the same thing but using a regression model

```{r}
MU3 = MU2[!(MU2$education_1=="#NULL!"),]

MU3 = MU3[!(MU3$age_1=="#NULL!"),]
MU3$age_1 = as.numeric(as.factor(MU3$age_1))

summary(glm(autism ~ binary + age_1 + education, data = MU3, family = "binomial"))
summary(glm(autism ~ as.factor(sex_1) + age_1 + education, data = MU3, family = "binomial"))
summary(glm(autism ~ relevel(as.factor(sex2), ref = "nonbinary") + age_1 + education, data = MU3, family = "binomial"))


```



Create some plots
```{R, Plot}

ggplot(MU2, aes(x=as.numeric(as.character(age_1)), colour=sex2)) +   geom_density() + theme_classic() + xlab("Age")

summary(aov(as.numeric(as.character(age_1)) ~ sex2, data = MU2))

summary_edu = setDT(MU2)[, list(count=.N) , list(sex2, education)]

summary_edu


ggplot(MU2, aes(sex2, fill = education)) + geom_bar(position = 'fill') + theme_classic()


```


